---
title: "FLRCpointFinder"
author: "Jean-Luc Jannink"
date: "6/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages  
```{r load packages, message=FALSE}
ip <- installed.packages()
packages_used <- c("tidyverse", "workflowr", "here", "lme4")
for (package in packages_used){
  if (!(package %in% ip[,"Package"])) install.packages(package)
}
library(tidyverse)

here::i_am("analysis/FLRCpointFinder.Rmd")

random_seed <- 45678
set.seed(random_seed)

key <- function(tbl, var, key){
  return(tbl[tbl %>% select(any_of(var)) == key,])
}
```

### Get data from the webscorer API
```{r Webscorer API}
courseData <- jsonlite::fromJSON(here::here("data", "CourseInfo.txt"))$Courses
courseData <- courseData %>% dplyr::mutate(Type=if_else(Type == "Mixed", "Road", Type))
apiURL <- "https://api.webscorer.com/racetimer/webscorerapi/results?raceid="
activity <- tibble()
for (rID in courseData$ID){
  rawWS <- jsonlite::fromJSON(paste0(apiURL, rID))$Racers
  rawWS <- rawWS %>% dplyr::mutate(Course=courseData$Name[courseData$ID==rID])
  activity <- activity %>% bind_rows(as_tibble(rawWS) %>% 
                                   select(Course, Name, RaceTime))
}
activity$Name[activity$Name == "akp84@cornell.edu"] <- "Adam Pacheck"
activity$Name[activity$Name == "adam.schoene@gmail.com"] <- "Adam Schoene"
```

### Curate the times
I'm using here an arbitrary cutoff of -5 for the studentized residual, but
eyeballing the outcome of the curation makes me believe that's right.  It
gets rid of times that are way too fast
```{r Analyse times}
activity <- activity %>% filter(RaceTime != 0) %>% dplyr::mutate(logTS=log(RaceTime))
fm <- lme4::lmer(logTS ~ Course + (1 | Name), data=activity)
activity <- activity %>% dplyr::mutate(stdntRes=resid(fm) / sigma(fm)) %>% 
  filter(stdntRes > -5)
fm <- lme4::lmer(logTS ~ Course + (1 | Name), data=activity)
activity <- activity %>% dplyr::mutate(stdntRes=resid(fm) / sigma(fm)) 
print(activity %>% filter(abs(stdntRes) > 3))
```

### Figure out the course time
For each course by athlete combination:  
1. If there are >1 times, take off the slowest
2. If there are >3 times, take off the fastest
3. Run the analysis with course as a fixed effect
```{r Trimmed times analysis}
athCrs <- activity %>% select(Name, Course) %>% distinct
trimmed <- function(ac){
  oneAC <- activity %>% filter(Name==unlist(ac[1]) & Course==unlist(ac[2])) %>% 
    arrange(logTS)
  nTimes <- nrow(oneAC)
  if (nTimes > 1) oneAC <- oneAC[-nTimes,]
  if (nTimes > 3) oneAC <- oneAC[-1,]
  return(oneAC)
}
acList <- apply(athCrs, 1, trimmed)
trimAC <- tibble()
dummy <- sapply(acList, function(tbl) trimAC <<- bind_rows(trimAC, tbl))
rm(dummy)
fm <- lme4::lmer(logTS ~ -1 + Course + (1 | Name), data=trimAC)
courseCov <- scale(lme4::fixef(fm), center=T, scale=F)
rownames(courseCov) <- substring(rownames(courseCov), 7)
courseData <- courseData %>% dplyr::mutate(courseTime = courseCov[Name, 1])
```

### Find fastest times per person per course
```{r Fastest times}
athCrs <- activity %>% select(Name, Course) %>% distinct
fastest <- function(ac){
  oneAC <- activity %>% filter(Name==unlist(ac[1]) & Course==unlist(ac[2]))
  return((oneAC %>% filter(RaceTime==min(RaceTime)))[1,])
}
acList <- apply(athCrs, 1, fastest)
fastestAC <- tibble(course=sapply(acList, function(tbl) unlist(tbl[1])),
                    athlete=sapply(acList, function(tbl) unlist(tbl[2])),
                    timeSec=sapply(acList, function(tbl) unlist(tbl[3]))
                    )
fastestAC <- fastestAC %>% filter(timeSec != 0)
```

### Analyse the times
```{r Analyse times}
fastestAC <- fastestAC %>% dplyr::mutate(logTS=log(timeSec),
  courseType=sapply(fastestAC$course, function(k) key(courseData, "Name", k)$Type),
  courseTime=sapply(fastestAC$course, function(k) key(courseData, "Name", k)$courseTime))
trimAC <- trimAC %>% dplyr::mutate(
  courseType=sapply(trimAC$Course, function(k) key(courseData, "Name", k)$Type),
  courseTime=sapply(trimAC$Course, function(k) key(courseData, "Name", k)$courseTime))
fm_ftp <- lme4::lmer(logTS ~ course + 
                   (1 + courseType | athlete), data=fastestAC)
ran_ftp <- lme4::ranef(fm_ftp)[[1]]
ran_ftp$roadStrength <- round(rank(ran_ftp$courseTypeTrail)/nrow(ran_ftp)*100, 1)
fm_ttptm <- lme4::lmer(logTS ~ Course + 
                   (1 + courseType + courseTime | Name), data=trimAC)
ran_ttptm <- lme4::ranef(fm_ttptm)[[1]]
ran_ttptm$speedStrength <- round(rank(ran_ttptm$courseTime)/nrow(ran_ttptm)*100, 1)
fm <- lme4::lmer(logTS ~ course + 
                   (1 | athlete), data=fastestAC)
fastestAC <- fastestAC %>% dplyr::mutate(deviation=resid(fm)) 
```

### Take a name, give a table
```{r Name to table}
name="Jean-Luc Jannink"
devTab <- fastestAC %>% filter(athlete == name) %>% arrange(timeSec) %>% 
  dplyr::mutate(devPercent=round(deviation*100, 1)) %>% 
  select(course, athlete, timeSec, devPercent) %>% 
  dplyr::mutate(targetSec=-round(timeSec*devPercent/100, 0)) %>% 
  dplyr::mutate(targetSec=if_else(targetSec < 0, targetSec, 0))
strength <- rank(ran_ftp$courseTypeTrail)
print(devTab)
```

